---
title: "Module2_Exercise2"
author: "Laura Huber"
date: "May 16, 2019"
output: html_document
---
## Solving differential equations revisited
```{r function}
require(deSolve)                          #deSolve library needed for this computing session

sir.model.open <- function (t, x, params) {    #here we begin a function with three arguments
  S <- x[1]                               #create local variable S, the first element of x
  I <- x[2]                               #create local variable I
  R <- x[3]                               #create local variable R
  with(                                   #we can simplify code using "with"
       as.list(params),                   #this argument to "with" lets us use the variable names
       {                                  #the system of rate equations
         dS <- mu*(S+I+R) - beta*S*I - mu*S
         dI <- beta*S*I - gamma*I - mu*I
         dR <- gamma*I - mu*R
         dx <- c(dS,dI,dR)                #combine results into a single vector dx
         list(dx)                         #return result as a list
       }
       )
}

```

### Specifying parameters
```{r parameters}
R0 <- 10
N <-  1					                       #population size
mu <- 0.02                             #per capita birth/death rate
gamma <- 365/10  			                 #recovery rate (in years)
beta <- R0*(gamma+mu)/N 	             #transmission rate
xstart <- c(S=0.2, I=0.001, R=1-0.2-0.001)	 #initial conditions, must sum to one
Tmax <- 120                            #integrate for 200 years after transients
params <- c(beta=beta, gamma=gamma, mu=mu)           #parameter vector
tau <- 0.1                             #size of time step
times <- seq(0, Tmax, by=tau)          #function seq returns a sequence

```

### Saving file
```{r saving file}
out <- ode(xstart,times,sir.model.open,params, method='ode45', rtol=1e-7)
```

### Plotting
```{r plotting}
op <- par(fig=c(0,0.5,0,1),mar=c(4,4,1,1))                      #set graphical parameters
plot(I~time,data=out, type='l', lwd=2)                          #plot the I variable against time
par(fig=c(0.5,1,0,1),mar=c(4,1,1,1),new=T)                      #re-set graphical parameters
plot(I~S,data=out,log='xy',yaxt='n',xlab='S', type='l', lwd=2)  #plot phase portrait
par(op) 
```

### Re-setting the parameters to t50
```{r re-setting}
xstart <- out[which(out[,1]==50),2:4]
```

## Adding pulsed vaccination

### Defining additional parameters
```{r additional parameters}
pv <- 0.1                     # fraction of susceptibles vaccinated
Tv <- 4                       # number of years between pulses
vacc.events <- floor(Tmax/Tv) # number of pulses in Tmax years
```

### Make a dataframe
```{r making dataframe}
data <- data.frame(S=out[which(out[,1]==50),2],
                   I=out[which(out[,1]==50),3],
                   R=out[which(out[,1]==50),4])
```

### for loop to add parameters
```{r for loop}
for(i in 1:vacc.events){
  out <- ode(xstart, seq(tau, Tv, by=tau), sir.model.open, params, method='ode45', rtol=1e-7)
  xstart <- out[dim(out)[1],2:4]        # reset initial condition
  xstart[1] <- (1-pv)*(tail(out,1)[2])  # vaccinate susceptibles
  xstart[3] <- xstart[3]+(pv)*(tail(out,1)[2])  # move to recovered class
  data <- rbind(data,out[,2:4])         # store result
}
```

### Plotting with the vaccinations

```{r plotting2}
data$time <- seq(50, Tmax+50, by=tau)
par(mar=c(5,4,4,4)+0.1)
plot(data$time[1:500], data$I[1:500], type='l', xlab='Time', ylab='', col='red', axes=FALSE)
axis(2, col.axis='red')
mtext(side=2, line=2.5, 'Infected', col='red')
box()
axis(1)
par(new=TRUE)
plot(data$time[1:500], data$S[1:500], type='l', xlab='', ylab='', axes=FALSE, col='black')
axis(4)
mtext(side=4, line=2.5, 'Susceptibles')
```


### Exercise 1: Why might this periodicity be of interest? What might be the consequences of "peaks" and "troughs" for public health?
In some cases, it might be beneficial to use pulsed vaccines, since it is sometimes more eficient logistically and more economic. However, avoiding picks is ideal. The objective is to find a pulse vaccine period that avoids peaks as much as possible.

### Exercise 2: By modifying the value of pv, see if you can locate the vaccination threshold. Does this agree with the analytically predicted threshold?
```{r}
#deSolve library needed for this computing session
ex2 <- function(.pvs){
sir.model.open <- function (t, x, params) {    #here we begin a function with three arguments
  S <- x[1]                               #create local variable S, the first element of x
  I <- x[2]                               #create local variable I
  R <- x[3]                               #create local variable R
  with(                                   #we can simplify code using "with"
       as.list(params),                   #this argument to "with" lets us use the variable names
       {                                  #the system of rate equations
         dS <- mu*(S+I+R) - beta*S*I - mu*S
         dI <- beta*S*I - gamma*I - mu*I
         dR <- gamma*I - mu*R
         dx <- c(dS,dI,dR)                #combine results into a single vector dx
         list(dx)                         #return result as a list
       }
       )
}


###################################################
### code chunk number 2: parameters
###################################################
R0 <- 10
N <-  1					                       #population size
mu <- 0.02                             #per capita birth/death rate
gamma <- 365/10  			                 #recovery rate (in years)
beta <- R0*(gamma+mu)/N 	             #transmission rate
xstart <- c(S=0.2, I=0.001, R=1-0.2-0.001)	 #initial conditions, must sum to one
Tmax <- 120                            #integrate for 200 years after transients
params <- c(beta=beta, gamma=gamma, mu=mu)           #parameter vector
tau <- 0.1                             #size of time step
times <- seq(0, Tmax, by=tau)          #function seq returns a sequence


###################################################
### code chunk number 3: pulsed-vaccination.rnw:113-114
###################################################
out <- ode(xstart,times,sir.model.open,params, method='ode45', rtol=1e-7)  


###################################################
### code chunk number 4: pulsed-vaccination.rnw:119-124
###################################################
# op <- par(fig=c(0,0.5,0,1),mar=c(4,4,1,1))                      #set graphical parameters
# plot(I~time,data=out, type='l', lwd=2)                          #plot the I variable against time
# par(fig=c(0.5,1,0,1),mar=c(4,1,1,1),new=T)                      #re-set graphical parameters
# plot(I~S,data=out,log='xy',yaxt='n',xlab='S', type='l', lwd=2)  #plot phase portrait
# par(op)                                                         #re-set graphical parameters


###################################################
### code chunk number 5: reset-parameters
###################################################
xstart <- out[which(out[,1]==50),2:4]


###################################################
### code chunk number 6: new-parameters
###################################################
pv <- .pvs                     # fraction of susceptibles vaccomated
Tv <- 4                       # number of years between pulses
vacc.events <- floor(Tmax/Tv) # number of pulses in Tmax years


###################################################
### code chunk number 7: dataframe
###################################################
data <- data.frame(S=out[which(out[,1]==50),2],
                   I=out[which(out[,1]==50),3],
                   R=out[which(out[,1]==50),4])


###################################################
### code chunk number 8: pulsed-vaccination.rnw:155-162
###################################################
for(i in 1:vacc.events){
  out <- ode(xstart, seq(tau, Tv, by=tau), sir.model.open, params, method='ode45', rtol=1e-7)
  xstart <- out[dim(out)[1],2:4]        # reset initial condition
  xstart[1] <- (1-pv)*(tail(out,1)[2])  # vaccinate susceptibles
  xstart[3] <- xstart[3]+(pv)*(tail(out,1)[2])  # move to recovered class
  data <- rbind(data,out[,2:4])         # store result
}


###################################################
### code chunk number 9: plot-vax
###################################################
data$time <- seq(50, Tmax+50, by=tau)
par(mar=c(5,4,4,4)+0.1)
plot(data$time[1:500], data$I[1:500], type='l', xlab='Time', ylab='', col='red', axes=FALSE)
axis(2, col.axis='red')
mtext(side=2, line=2.5, 'Infected', col='red')
box()
axis(1)
par(new=TRUE)
plot(data$time[1:500], data$S[1:500], type='l', xlab='', ylab='', axes=FALSE, col='black')
axis(4)
mtext(side=4, line=2.5, 'Susceptibles')

}



sample.pvs <- seq(0.1,0.95,by=0.05)
for(i in 1:length(sample.pvs)){
  ex2(sample.pvs[i])
}


```

When anlyzing a range of pv values from 0.1 to 0.95, by 0.05, we can see that when we have a pv of around 0.55, the Time coincides with t=50, that we had previously observed, that caused the infected numbers to be close to 0 and erradicate the disease.

## Exercise 3: One thing we might be interested in knowing is how our vaccination strategy affects mean disease prevalance. Devise a strategy to study this question and produce a plot illustrating the result."

When using the means of infectious numbers across the pv values, we can see that when pv calues is around 0.55, the prevalence is close to 0.

```{r code for prevalence}
ex2 <- function(.pvs){
sir.model.open <- function (t, x, params) {    #here we begin a function with three arguments
  S <- x[1]                               #create local variable S, the first element of x
  I <- x[2]                               #create local variable I
  R <- x[3]                               #create local variable R
  with(                                   #we can simplify code using "with"
       as.list(params),                   #this argument to "with" lets us use the variable names
       {                                  #the system of rate equations
         dS <- mu*(S+I+R) - beta*S*I - mu*S
         dI <- beta*S*I - gamma*I - mu*I
         dR <- gamma*I - mu*R
         dx <- c(dS,dI,dR)                #combine results into a single vector dx
         list(dx)                         #return result as a list
       }
       )
}


###################################################
### code chunk number 2: parameters
###################################################
R0 <- 10
N <-  1					                       #population size
mu <- 0.02                             #per capita birth/death rate
gamma <- 365/10  			                 #recovery rate (in years)
beta <- R0*(gamma+mu)/N 	             #transmission rate
xstart <- c(S=0.2, I=0.001, R=1-0.2-0.001)	 #initial conditions, must sum to one
Tmax <- 120                            #integrate for 200 years after transients
params <- c(beta=beta, gamma=gamma, mu=mu)           #parameter vector
tau <- 0.1                             #size of time step
times <- seq(0, Tmax, by=tau)          #function seq returns a sequence


###################################################
### code chunk number 3: pulsed-vaccination.rnw:113-114
###################################################
out <- ode(xstart,times,sir.model.open,params, method='ode45', rtol=1e-7)  


###################################################
### code chunk number 4: pulsed-vaccination.rnw:119-124
###################################################
# op <- par(fig=c(0,0.5,0,1),mar=c(4,4,1,1))                      #set graphical parameters
# plot(I~time,data=out, type='l', lwd=2)                          #plot the I variable against time
# par(fig=c(0.5,1,0,1),mar=c(4,1,1,1),new=T)                      #re-set graphical parameters
# plot(I~S,data=out,log='xy',yaxt='n',xlab='S', type='l', lwd=2)  #plot phase portrait
# par(op)                                                         #re-set graphical parameters


###################################################
### code chunk number 5: reset-parameters
###################################################
xstart <- out[which(out[,1]==50),2:4]


###################################################
### code chunk number 6: new-parameters
###################################################
pv <- .pvs                     # fraction of susceptibles vaccomated
Tv <- 4                       # number of years between pulses
vacc.events <- floor(Tmax/Tv) # number of pulses in Tmax years


###################################################
### code chunk number 7: dataframe
###################################################
data <- data.frame(S=out[which(out[,1]==50),2],
                   I=out[which(out[,1]==50),3],
                   R=out[which(out[,1]==50),4])


###################################################
### code chunk number 8: pulsed-vaccination.rnw:155-162
###################################################
for(i in 1:vacc.events){
  out <- ode(xstart, seq(tau, Tv, by=tau), sir.model.open, params, method='ode45', rtol=1e-7)
  xstart <- out[dim(out)[1],2:4]        # reset initial condition
  xstart[1] <- (1-pv)*(tail(out,1)[2])  # vaccinate susceptibles
  xstart[3] <- xstart[3]+(pv)*(tail(out,1)[2])  # move to recovered class
  data <- rbind(data,out[,2:4])         # store result
}


###################################################
### code chunk number 9: plot-vax
###################################################
data$time <- seq(50, Tmax+50, by=tau)
# par(mar=c(5,4,4,4)+0.1)
# plot(data$time[1:500], data$I[1:500], type='l', xlab='Time', ylab='', col='red', axes=FALSE)
# axis(2, col.axis='red')
# mtext(side=2, line=2.5, 'Infected', col='red')
# box()
# axis(1)
# par(new=TRUE)
# plot(data$time[1:500], data$S[1:500], type='l', xlab='', ylab='', axes=FALSE, col='black')
# axis(4)
# mtext(side=4, line=2.5, 'Susceptibles')

meanPrevalence <- mean(data$I)
}

meanPrevs <- c()

sample.pvs <- seq(0.1,0.95,by=0.05)
for(i in 1:length(sample.pvs)){
  meanPrevs[i] <- ex2(sample.pvs[i])
}
plot(sample.pvs, meanPrevs)

```








